---
title: "Stat 3302 Project"
author: "Molly Hunold"
date: "4/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Loading Data and Libraries, message = FALSE, error = FALSE}
library(tidyverse)
library(lubridate)
library(kableExtra)
library(broom)
players  = read_csv("C:/Users/mehun/Documents/8 - Spring 2021/Stat 3302 Statistical Modeling for Discovery II/Homework/data.csv")
```

## Data Transformation

The date of birth is currently a character string like "MM/DD/YY", I want to convert this to a single number 1-365 for the players date of birth, regardless of year. 

```{r Convert DOB}
# changed to %Y instead of %y b/c yr given w/ century. was producing two NA values.
players$dateOfBirth = as.Date.character(players$dateOfBirth, format = "%m/%d/%Y")
players$dateOfBirth = yday(players$dateOfBirth)
players$wb = ceiling(players$dateOfBirth / 7)
##for now, ignoring players in week 53
#players = players %>% filter(players$wb < 53)
# change so that wk 53 is included in wk 52
players$wb[players$wb==53] <- 52
players$tb = (players$wb - 0.5)/52
```

## Table 1

I haven't adjusted my data yet to correct for the 53rd birthweek like they did in the paper, but the t tests loosly match (i'd imagine once the 53rd birthweek is adjusted they'd match exactly). 

Recreating the t-test of mean section of Table 1

```{r}
# get weighted mean for all players. this was just me checking if it matched.
#weights <- if_else(players$wb==52, 7/8.25, 1)
#weighted.mean(players$tb, weights)

countrylist = c("Argentina", "Netherlands", "Turkey", "Belgium", 
                "Spain", "Italy", "Germany", "Portugal", "Brazil", 
                "Russia", "France")
grouped_data = players %>% 
  group_by(country) %>%
  filter(country == "Argentina" | country == "Netherlands" | country == "Turkey" |
         country == "Belgium"  | country == "Spain" | country ==  "Italy" |  country == "Germany"  |
         country == "Portugal" | country == "Brazil" | country == "Russia" | country == "France") %>%
  summarize(n = n(),
            #xbar = mean(tb),
            xbar = weighted.mean(tb, if_else(wb==52, 7/8.25, 1)),
            s =  sd(tb))
grouped_data$t = (grouped_data$xbar - 0.5) / (grouped_data$s / sqrt(grouped_data$n))
grouped_data$p = pt(grouped_data$t,  df  =  grouped_data$n)
grouped_data$p = if_else(grouped_data$p < 0.001, "< .001", if_else(grouped_data$p < 0.01, "< 0.01", if_else(grouped_data$p < 0.05, "< 0.05", " ")))
grouped_data$tb = round(grouped_data$xbar, 3)
grouped_data = grouped_data %>% select(country,n,tb,p) %>% arrange(tb)
grouped_data %>% kable()
```

Recreating the poisson regression - this all loosely fits for the model for all countries.

```{r}
#Need the data as Team, Week, N
players = as_tibble(players)
regression_data = players %>%
  group_by(wb) %>%
  summarize(n = n()) %>%
  mutate(tb = (wb-0.5)/52)
#Deflate frequency of wk 52 by 7/8.25. R gives a warning bc that count is no longer an integer.
regression_data$n[regression_data$wb==52] <- regression_data$n[regression_data$wb==52] * (7/8.25)
regression_data %>% kable()
#This is the model for all countries
(pois_model = glm(n~tb, data  = regression_data, family = "poisson"))
(r2 = 1- (pois_model$deviance/pois_model$null.deviance))
```

But I havent figured out how to match the models for individual countries...

```{r}
#This is the model for just one country?
regression_data = players %>%
  filter(country == "Argentina") %>%
  group_by(wb) %>%
  summarize(n = n()) %>%
  mutate(tb = (wb-0.5)/52)
#Deflate frequency of wk 52 by 7/8.25. R gives a warning bc that count is no longer an integer.
regression_data$n[regression_data$wb==52] <- regression_data$n[regression_data$wb==52] * (7/8.25)
pois_model = glm(n~tb, data = regression_data, family = "poisson")
summary(pois_model)
```